Reproducible Research Course Project 1
================

Load the Data



```{r}
# Code for reading the data 
library(knitr)
getwd()
setwd("C:/users/robert/documents/data/ReproducibleResearch/CourseProject2/repdata_data_activity")
activity<-read.csv("activity.csv")
```


The total number of steps per day 
```{r}

##The total number of steps taken each day
bydaySum<-(aggregate(steps~date,data=activity,FUN=sum))
bydaySum
```


A histogram of the total number of steps per day  
```{R hist, fig.weidth=-8, fig.height=6}
##Histogram of total number of steps taken each day 
hist(bydaySum$steps, breaks=10, col="red", ylim = c(0,20), main="Average of Total Daily Steps", xlab="Steps")
```



The mean and median of the total number of steps taken per day 
```{r}
#the mean and median of the total number of steps taken per day 
mean(bydaySum$steps)
median(bydaySum$steps)
```

The average daily activity pattern 
```{R}

#plot total of daily average number of steps taken per interval 
int<-aggregate(steps~interval,data=activity, FUN=mean)
intPlot<-with(int,plot(interval, steps, type="l", main="Average Daily Activity Pattern", col="red", lwd=2))
```

The five-minute interval that contains the maximum number of steps
```{R}
int[which.max(int$steps), ]
```

Imputing Missing Values 

The total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{R}
sum(is.na(activity))
```

Create a new dataset that fills in missing values using mean for each day 
```{R}
#create a new dataset with the NA values filled in 
activity2<-activity
naValues<-is.na(activity2$steps)
interval_mean<-tapply(activity2$steps, activity2$interval, mean, na.rm=TRUE, simplify=TRUE)
activity2$steps[naValues]<-interval_mean[as.character(activity2$interval[naValues])]
names(activity2)
sum(is.na(activity2))

```

Histogram of the total number of step taken each day 
```{R}
bydaySum2<-(aggregate(steps~date,data=activity2,FUN=sum))
hist(bydaySum2$steps, breaks=15, xlab="Average Total Steps by Day", main="Histogram of Average Total Steps with Missing Values per Day", ylim=c(0, 25), col="red")
```

The mean and median total number of steps taken per day 

```{R}
mean(bydaySum2$steps)
median(bydaySum2$steps)
```

The mean and median do not differ.  The histogram is different.   

The impact of imputing missing data on the estimates of the total daily number of steps? 
```{R}
summary(bydaySum)
summary(bydaySum2)
```

Are there any difference in activity patterns between weekdays and weekends? 

```{R}
#create a new factor variable in the dataset with two levels-"weekday" and "weekend"
activity2$date<-as.Date(activity2$date)  
weekdays1<-c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

activity2$weekDay<-factor((weekdays(activity2$date) %in% weekdays1),levels=c(FALSE, TRUE), labels = c("Weekend", "Weekday"))
```


A panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekdays or weekenend days 
```{R}
library(lattice)
int2<-aggregate(steps~weekDay+interval,data=activity2, FUN=mean)
xyplot(steps~interval|weekDay, data=int2, layout=c(1,2), type="l", col="red")
```





